---
title: 'Lab week 5: time series and a map!'
author: "Mireia Valle"
date: "6/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### Attach packages
```{r}
#Packages for general stuff:
library(tidyverse)
library(janitor)
library(lubridate)#change dates
library(here)
#install.packages("paletteer")
library(paletteer)#color palettees

#Packages for time series stuff:
#install.packages("tsibble")
library(tsibble)
#install.packages("fable")
library(fable)
#install.packages("fabletools")
library(fabletools)
#install.packages("feasts")
library(feasts)
#install.packages("forecast")
library(forecast)

#Packages for spatial stuff:
library(sf)
library(tmap)
library(mapview)
```


### Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
                       clean_names()
```

-Make description all lowercase
-only keep observation where "description" variable contains "production"
-remove any observation where "description"variable contains "total"

```{r}
renew_clean <- us_renew %>%
mutate(description = str_to_lower(description)) %>%
filter(str_detect(description, pattern = "consumption")) %>%
filter(!str_detect(description, pattern = "total"))
```

### Convert "yyyymm" column to a date

```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

class(renew_date$month_sep)
```

# Make a version where I have the month & year in a separate column 

```{r}
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

## Look at it:

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value, 
                                          group = description)) +
  geom_line(aes(color = description))

renew_gg
```

## Looking at colors from "paleteer" pckg

View(palettes_d_names)

Updating my colors with paletteer palettes:
```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::bigsur") #we use _d because we want discrete colors

```

```{r}
class(renew_parsed)
```


### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a couple of different ways

```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
renew_ts %>% gg_season (value)

#season plot in ggplot

ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~ description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
```

### Just look at the hydroelectric energy consumption

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))
```

### What if I want quarterly average consumption for hydro?

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose that hydro_ts (look at the different scales)

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()
```



```{r}
hist(components(dcmp)$remainder)
```

Ploting 
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

# DANGER DANGER
```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ETS(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

#ploting the forecasted data
hydro_model %>% autoplot()

#ploting the data + forecasted data
hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))

```

### Make a world map!

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

